---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
df = read.csv('activity.csv',as.is=T)
```


## What is mean total number of steps taken per day?

```{r total-day-histo,dpi=100}
df.no.NA = subset(df,!is.na(steps))
per.day = split(df.no.NA,df.no.NA$date)
per.day.total = sapply(per.day,function(x) sum(x$steps))
hist(
        per.day.total,
        xlab = 'Total number of steps taken each day',
        las = 1,main=''
        )
mean = mean(per.day.total)
median = median(per.day.total)
```
The mean is `r mean` and the median is `r median` days.


## What is the average daily activity pattern?

```{r activity-pattern,dpi=100}
per.interval = split(df.no.NA,df.no.NA$interval)
per.interval.avg = sapply(per.interval,function(x) mean(x$steps))
plot(
        x = names(per.interval.avg),xlab='5-minute interval',
        y = per.interval.avg,ylab='Average number of steps taken',
        las = 1, type='l',bty='n'
        )

m = which.max(per.interval.avg)
```

The 5-minute interval `r names(m)` contains the maximum number of steps, on average across all the days in the dataset: `r per.interval.avg[m]`.



## Imputing missing values

```{r}
na = !complete.cases(df)
table(na)
```
`r sum(na)` lines contains missing values.

We impute the missing values using the mean for that 5-minute interval.
```{r}
withNA = df[!complete.cases(df),]
withNA$steps = per.interval.avg[as.character(withNA$interval)]
dfImputed = rbind(df.no.NA,withNA)
```

```{r total-day-histo-imputed,dpi=100}
per.day = split(dfImputed,dfImputed$date)
per.day.total = sapply(per.day,function(x) sum(x$steps))
hist(
        per.day.total,
        xlab = 'Total number of steps taken each day',
        las = 1,main=''
        )
mean2 = mean(per.day.total)
median2 = median(per.day.total)
```
The mean is `r mean` and the median is `r median` days.

What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
mean==mean2
diff = median - median2
```
It shifts (increases) the median by `r abs(diff)`.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
date = as.Date(dfImputed$date)
dfImputed$day = weekdays(date)
dfImputed$we = factor(
        dfImputed$day %in% c('Saturday','Sunday'),
        labels = c('weekend','weekday'),
        levels = c('TRUE','FALSE')
        )
```


```{r panel-plot,dpi=100}
library('lattice')
dfImputedAvg = unique(dfImputed[,c('interval','we')])
dfImputedAvg$steps = NA
for (i in 1:nrow(dfImputedAvg)){
        s = subset(dfImputed,
                interval==dfImputedAvg$interval[[i]]&
                        we==dfImputedAvg$we[[i]])
        dfImputedAvg$steps[[i]] = mean(s$steps)
}
xyplot(steps ~ interval | we,data=dfImputedAvg,type='l')
```

There are indeed some clear differences between weekend and weekdays, where the number of steps is more constant over the day during the weekend.

```{r}
sessionInfo()
```

